---
title: "League of Dionysus 2020 Stats"
author: "Noel Peterson"
date: "`r format(Sys.time(), '%x')`"
output: html_document
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), 'index.html')) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  fig.retina = 4,
  fig.align = "center",
  out.width = "100%"
)

library(tidyverse)
library(see)
```

```{r load-data, message=FALSE}
# Read weekly scores and calculative cumulative score
SCORES <- read_csv("./weekly_scores.csv") %>%
  as_tibble() %>%
  pivot_longer(cols=starts_with("Week"),
               names_to="Week",
               names_prefix="Week",
               names_transform=list(Week=as.integer),
               values_to="Score") %>%
  group_by(Week) %>%
  arrange(Week, Score) %>%
  mutate(
    Winner = if_else(Score == max(Score), 1, 0),
    Rank = 8 - rank(Score, ties.method="max")
  ) %>%
  group_by(Team) %>%
  arrange(Team, Week) %>%
  mutate(
    CumulativeScore = cumsum(Score),
    CumulativeMeanScore = cummean(Score),
    CumulativeWins = cumsum(Winner),
    CumulativeWinnings = CumulativeWins * 9.24
  ) %>%
  group_by(Week) %>%
  arrange(Week, CumulativeScore) %>%
  mutate(CumulativeRank = 8 - rank(CumulativeScore, ties.method="max")) %>%
  ungroup() %>%
  arrange(Team, Week) %>%
  mutate(Team = as_factor(Team))
```


## Standings and stats
#### Through Week `r max(SCORES$Week)`

```{r summary}
# Calculate summaries
SUMMARY <- SCORES %>%
  group_by(Team) %>%
  summarize(
    TotalScore = max(CumulativeScore),
    MinScore = min(Score),
    MaxScore = max(Score),
    MeanScore = mean(Score),
    SDScore = sd(Score),
    MeanRank = mean(Rank),
    SDRank = sd(Rank),
    WeeklyWins = max(CumulativeWins),
    Winnings = max(CumulativeWinnings),
    .groups="drop"
  )

FIRST_WIN <- SCORES %>%
  filter(CumulativeWins == 1, Winner == 1) %>%
  select(Team, Week) %>%
  rename(FirstWinWeek = Week)

SUMMARY <- left_join(SUMMARY, FIRST_WIN, by="Team")

SUMMARY %>%
  arrange(desc(TotalScore)) %>%
  transmute(
    `Rank` = 8 - rank(TotalScore, ties.method="max"),
    `Team` = Team,
    `Total Score` = TotalScore,
    `Wins` = WeeklyWins,
    `Winnings` = sprintf("$%.2f", Winnings),
    `Average Score` = round(MeanScore, 2),
    `Low Score` = MinScore,
    `High Score` = MaxScore
  ) %>%
  knitr::kable(align='clrcrrrr')

MEAN_WIN_SCORE <- SCORES %>% filter(Winner == 1) %>% .$Score %>% mean() %>% round(2)
```

The average weekly winning score has been **`r MEAN_WIN_SCORE`** points.

<hr>

## A series of charts

```{r weekly-scores-line}
# Plot weekly scores
ggplot(SCORES, aes(x=Week, y=Score, color=Team)) +
  geom_line(size=1) +
  scale_x_continuous(breaks=1:17, expand=expansion(add=0)) +
  scale_y_continuous(limits=c(0, max(SCORES$Score)+10), expand=expansion(add=0)) +
  scale_color_see() +
  theme_modern() +
  labs(title="Weekly Scores", y=NULL)

# # Plot weekly ranks
# ggplot(SCORES, aes(x=Week, y=(-1*Rank), color=Team)) +
#   geom_line(size=1) +
#   scale_y_continuous(breaks=-1:-7, labels=1:7) +
#   scale_x_continuous(breaks=1:17, expand=expansion(add=0)) +
#   scale_color_see() +
#   theme_modern() +
#   labs(title="Weekly Finishes", y=NULL)
```

<hr style="width: 80%; height: 1px; background-image: linear-gradient(to right, rgba(0,0,0,0), rgba(0,0,0,0.25), rgba(0,0,0,0));">

```{r weekly-scores-heat}
# Plot heat map of weekly scores
left_join(SCORES, select(SUMMARY, Team, MeanScore), by="Team") %>%
  ggplot(., aes(x=Week, y=fct_reorder(Team, MeanScore), fill=Score)) +
    geom_raster() +
    scale_x_continuous(breaks=1:17, expand=expansion(add=0)) +
    scale_y_discrete(expand=expansion(add=0)) +
    scale_fill_distiller(palette="YlGn", direction=1) +
    theme_modern() +
    labs(title="Heat Map of Weekly Scores", y=NULL)
```

<hr style="width: 80%; height: 1px; background-image: linear-gradient(to right, rgba(0,0,0,0), rgba(0,0,0,0.25), rgba(0,0,0,0));">

```{r cumulative-scores}
# Plot cumulative scores
ggplot(SCORES, aes(x=Week, y=CumulativeScore, color=Team)) +
  geom_line(size=1) +
  scale_x_continuous(breaks=1:17, expand=expansion(add=0)) +
  scale_y_continuous(limits=c(0, max(SCORES$CumulativeScore)+10), expand=expansion(add=0),
                     labels=scales::label_comma()) +
  scale_color_see() +
  theme_modern() +
  labs(title="Total Scores Over Time", y=NULL)
```

<hr style="width: 80%; height: 1px; background-image: linear-gradient(to right, rgba(0,0,0,0), rgba(0,0,0,0.25), rgba(0,0,0,0));">

```{r cumulative-ranks}
# Plot cumulative ranks
ggplot(SCORES, aes(x=Week, y=(-1*CumulativeRank), color=Team)) +
  geom_line(size=1) +
  scale_y_continuous(breaks=-1:-7, labels=1:7) +
  scale_x_continuous(breaks=1:17, expand=expansion(add=0)) +
  scale_color_see() +
  theme_modern() +
  labs(title="Overall Standings Over Time", y=NULL)

# # Plot cumulative mean scores
# ggplot(SCORES, aes(x=Week, y=CumulativeMeanScore, color=Team)) +
#   geom_line(size=1) +
#   scale_x_continuous(breaks=1:17, expand=expansion(add=0)) +
#   scale_y_continuous(limits=c(0, max(SCORES$CumulativeMeanScore)+10),
#                      expand=expansion(add=0)) +
#   scale_color_see() +
#   theme_modern() +
#   labs(title="Mean Scores Over Time", y=NULL)
```

<hr style="width: 80%; height: 1px; background-image: linear-gradient(to right, rgba(0,0,0,0), rgba(0,0,0,0.25), rgba(0,0,0,0));">

```{r cumulative-winnings}
# Plot cumulative winnings
# left_join(SCORES, select(SUMMARY, Team, FirstWinWeek), by="Team") %>%
#   ggplot(., aes(x=Week, y=CumulativeWinnings,
#              fill=fct_reorder(Team, FirstWinWeek))) +
ggplot(SCORES, aes(x=Week, y=CumulativeWinnings, fill=Team)) +
    geom_col(width=0.95) +
    scale_x_continuous(breaks=1:17, expand=expansion(add=0)) +
    scale_y_continuous(limits=c(0, 10*max(SCORES$Week)), expand=expansion(add=0),
                       labels=scales::label_dollar()) +
    scale_fill_see() +
    theme_modern() +
    labs(title="Total Winnings Over Time", y=NULL) +
    guides(fill=guide_legend(title="Team"))
```

<hr style="width: 80%; height: 1px; background-image: linear-gradient(to right, rgba(0,0,0,0), rgba(0,0,0,0.25), rgba(0,0,0,0));">

```{r weekly-scores-freq}
# Plot frequency distribution of weekly scores
left_join(SCORES, select(SUMMARY, Team, MeanScore), by="Team") %>%
  ggplot(., aes(x=Score, fill=Team)) +
    geom_density(size=0) +
    scale_x_continuous(limits=c(min(SCORES$Score), max(SCORES$Score)),
                       breaks=seq(0, 300, 20),
                       expand=expansion(add=0)) +
    scale_y_continuous(expand=expansion(add=0)) +
    facet_grid(rows=vars(fct_reorder(Team, MeanScore, .desc=TRUE)), switch="y") +
    scale_fill_see() +
    theme_modern(legend.position="none") +
    labs(title="Frequency Distribution of Weekly Scores", y=NULL) +
    guides(fill=guide_legend(title="Team")) +
    theme(axis.line.y=element_blank(), axis.text.y=element_blank(),
          strip.text.y.left=element_text(size=13, angle=0,face="plain", hjust=1))

# left_join(SCORES, select(SUMMARY, Team, SDScore), by="Team") %>%
#   ggplot(., aes(x=Score, fill=fct_reorder(Team, SDScore, .desc=TRUE))) +
#   geom_density(position="stack", size=0) +
#   scale_fill_see()
#   theme_modern() +
#   labs(title="Frequency Distribution of Weekly Scores", y="Probability") +
#   guides(fill=guide_legend(title="Team"))
```

<hr>

## So you're tellin' me there's a chance...

```{r target}
# Estimate what's needed for each team to win the Championship
TARGET_SCORE <- SCORES %>%
  filter(CumulativeScore == max(CumulativeScore)) %>%
  left_join(SUMMARY, by="Team") %>%
  mutate(Target = CumulativeScore + MeanScore * (17 - Week)) %>%
  .$Target %>%
  round(2)
```

The leading team is on pace to end the season with **`r TARGET_SCORE`** total points. The following table shows what each team would need to score in each of the remaining games to match that score.

```{r catch-up}
CATCH_UP <- SCORES %>%
  filter(Week == max(Week)) %>%
  left_join(SUMMARY, by="Team") %>%
  select(Team, Week, CumulativeScore, MeanScore) %>%
  mutate(
    TargetMeanScore = (TARGET_SCORE - CumulativeScore) / (17 - Week),
    PercentIncrease = (TargetMeanScore / MeanScore - 1) * 100
  ) %>%
  arrange(desc(CumulativeScore))

CATCH_UP %>%
  transmute(
    `Team` = Team,
    `Average Score` = round(MeanScore, 2),
    `Target Score for Remaining Weeks` = round(TargetMeanScore, 2),
    `Percent Improvement Needed` = sprintf("%.2f%%", PercentIncrease)
  ) %>%
  knitr::kable(align="lrrr")
```

Good luck, everyone!
